Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations

### 2021-12-30 analysis/paper/paper.Rmd
TB results edits

We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish caught using hook-and-line angling. The CPUE, measured as number of Northern Pikeminnow caught per angler hour, was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)). During the mark-recapture efforts, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019 which included `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals (i.e., `r as.character(english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m))` individual was captured more than once). There were an additional `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020 which included `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals. Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.

Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)). In both cases, the multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)). Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture efforts.

All mark-recapture estimators yielded estimates with overlapping confidence intervals. However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day), we fell those estimators most appropriate. Additionally, because the Schnabel estimator includes a small sample size modification, as in our case (i.e., few recaptures), we feel it is most appropriate. The Schnabel estimator has the added benefit of more precise confidence intervals and is more straightforward than the Schumacher-Eschmeyer estimate as it is essentially the weighted average among multiple Chapman-modified Lincoln-Peterson estimates. Therefore, further analyses and results will be based on the Schnabel estimates of abundance.

Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across the two fall sampling events, `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`), and multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.

@@

```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events. Note the y-axis is truncated at 100."}
all_est %>%
  mutate(across(c(N, Lci, Uci),
                ~ . / 1000)) %>%
  ggplot(aes(x = model,
             y = N,
             color = model)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = Lci,
                    ymax = Uci),
                width = 0.1) +
  facet_wrap(~ event_name) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(x = "Estimator",
       y = "Abundance (x 1,000)") +
  #theme_bw() +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        strip.text.x = element_text(size = 11,
                                    color = "black")) +
  scale_color_brewer(palette = "Dark2")

>>>>

Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish. The CPUE, measured as number of Northern Pikeminnow caught per angler hour, was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)). During the mark-recapture efforts, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019 which included `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals. There were an additional `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020 consisting of `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals. Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.

Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)). In both cases, the multiple census estimators were larger than the single census estimators. Confidence intervals for the fall 2020 effort were more varied, likely resulting from fewer recaptures in that year.

All mark-recapture estimators yielded estimates with overlapping confidence intervals. However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day) and required a small sample size modification, we used the Schnabel estimates of abundance for further analyses and results. Accordingly, the mean of Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`). By multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.

<!--However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day), we fell those estimators most appropriate. Additionally, because the Schnabel estimator includes a small sample size modification, as in our case (i.e., few recaptures), we feel it is most appropriate. The Schnabel estimator has the added benefit of more precise confidence intervals and is more straightforward than the Schumacher-Eschmeyer estimate as it is essentially the weighted average among multiple Chapman-modified Lincoln-Peterson estimates. Therefore, further analyses and results will be based on the Schnabel estimates of abundance.-->

@@

```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events. Single census estimators are shown as circles; multiple census estimators are shown as triangles. Note the y-axis is truncated at 100 and the upper confidence interval for the Schumacker-Eschmeyer estimator in 2020 extends beyond that."}
all_est %>%
  mutate(across(c(N, Lci, Uci),
                ~ . / 1000)) %>%
  mutate(census = case_when(grepl("Chapman", model) ~ "single",
                            grepl("Petersen", model) ~ "single",
                            grepl("Schnabel", model) ~ "multiple",
                            grepl("Schumacher-Eschmeyer", model) ~ "multiple")) %>%
  ggplot(aes(x = model,
             y = N,
             #color = model
             )) +
  geom_point(aes(size = 4,
                 shape = census)) +
  scale_shape_manual(values = c(17, 19)) +
  geom_errorbar(aes(ymin = Lci,
                    ymax = Uci),
                width = 0.1) +
  facet_wrap(~ event_name) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(x = "Estimator",
       y = "Abundance (x 1,000)") +
  #theme_bw() +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        strip.text.x = element_text(size = 11,
                                    color = "black")) +
  scale_color_brewer(palette = "Dark2")
